---
title       : Wanna Lose Weight?
subtitle    : A calorie consumption calculator made with R and Shiny
author      : Edman P. Anjos
job         : 
framework   : io2012
highlighter : highlight.js
hitheme     : tomorrow
widgets     : [mathjax]
mode        : selfcontained
---

## Wanna lose Weight?

- Losing weight is easier than you imagine.
- The human body consumes an specific amount of calories per day.
- Consuming less calories than said amount will create an energy deficit.
- In order to make up for that, our body will fall back and use the energy
  previously stored as fat (for the harsh times).
- Do this a few days in a row and even your friends will see the difference.

---

## Then how many calories should I take?

- The number of calories consumed by your body depends on several factors,
  including:
  - Weight
  - Age
  - Height
  - Gender
  - Average amount of physical activities done throughout the week

- Consume more than this amount and you will gain weight
- Consume exactly this amount and you will maintain your weight
- Consume less than this amount and you will lose weight

---

## And how exactly does it work?

- It's just science: we apply the [Harris-Benedict
  formula](http://en.wikipedia.org/wiki/Harris%E2%80%93Benedict_equation).

- First we compute your basal metabolic rate ($BMR$), that is what your body
  consumes just for its natural functions (without any physical activity
  included).
  - For a man, the equation to calculate the $BMR$ is as follows
    $$BMR = 88.362 + (13.397 W) + (4.799 H) - (5.677 A)$$
  - Being $W$ the weight in kilograms, $H$ the height in centimeters and $A$
    the age in years.

- Then we use the Harris-Benedict principle to mix in the equation the
  calories consumed according to the amount of exercise you practice in a week.
  - For a person who practices little to no exercise, the daily calorie
    consumption is
	$$KCal = 1.2 BMR$$

---

## I'm almost convinced, now show me the numbers

- Let's calculate the daily calorie consumption of a 21-year-old, 73kg man
  who is 182cm tall and hits the gym some four days per week.

```{r}
bmr <- 88.362 + (13.397 * 73) + (4.799 * 182) - (5.677 * 21)
```
```{r, echo=FALSE}
bmr
```

- Our friend's BMR is ```r bmr ```. That's what his body consumes just for its
  natural functions.
- Now let's add in the equation the amount of exercise he practices per week

```{r}
kcal <-  1.55 * bmr
```
```{r, echo=FALSE}
kcal
```

- That's the number our guy needs to know: ```r kcal ``` is his actual daily
  calorie consumption. In order to lose weight he has to consume less than that
  everyday.

